The state of knowledge for marine invertebrate biodiversity in the continental US

### 2017-12-13 manuscript.Rmd
[fix #71] finish itis

## Results

@@

We retrieved `r format(n_distinct(combined_species$worms_valid_name), big.mark = ",")` species (`r n_distinct(na.omit(combined_species$worms_phylum))` phyla) of marine invertebrates in the continental US EEZ by combining  `r format(nrow(idigbio_records), big.mark=",")` museum records from iDigBio (`r format(n_distinct(idigbio_records$worms_valid_name), big.mark = ",")` species, `r n_distinct(na.omit(idigbio_records$worms_phylum))` phyla), and `r format(nrow(obis_records), big.mark=",")` records from OBIS (`r format(n_distinct(obis_records$worms_valid_name), big.mark=",")` species,
`r n_distinct(na.omit(obis_records$worms_phylum))` phyla). Extant marine invertebrates occurring in the US represent less than `r format_output(nrow(idigbio_records)/idigbio_stats$total_no_fossil*100)`% of the total number of non-fossil iDigBio records (`r round(idigbio_stats$total_no_fossil/1e6, 1)` million), and `r format_output(nrow(idigbio_records)/idigbio_stats$usa_no_fossil*100)`% of the US records (`r round(idigbio_stats$usa_no_fossil/1e6, 1)` million). Marine invertebrates that occur in the US represent 4.1% of the 48 million occurrence records in OBIS.

@@

Marine invertebrates are underrepresented in iDigBio and OBIS given their diversity. We compared the total number of species represented in iDigBio and OBIS for each major division of life to the total described diversity based on the WoRMS estimates[@worms-stats]. While marine vertebrates represent  `r total_vert_prop`% of the global diversity, they represent `r idigbio_vert_prop`% and `r obis_vert_prop`% of the species in iDigBio and OBIS respectively (Fig. \@ref(fig:kingdom-diversity-plot)). This underrepresentation is also apparent in the collecting effort: the median number of specimens in iDigBio per species ranges between `r kng_medians %>% filter(sub_kingdom=="animalia") %>% pull(rg)` for invertebrate phyla, but is `r kng_medians %>% filter(sub_kingdom=="animalia - vertebrates") %>% pull(rg)` for vertebrates.

```{r itis-stats}
prop_molluscs <- itis_molluscs_in_dbs %>%
    dplyr::summarize(
               prop_idigbio_us = format_output(mean(n_idigbio_geo_us > 0,
                                                    na.rm = TRUE)),
               prop_obis_us = format_output(mean(n_obis_geo_us > 0,
                                                 na.rm = TRUE))
           )

To assess the fraction of digitization of the fauna, we compared the species included in iDigBio and OBIS with the comprehensive checklists for molluscs [@turgean1998] and crustaceans[@mclaughlin2005] that are available for the United States. For molluscs, of the `r nrow(itis_molluscs_in_dbs)` species included in the checklist,

<!-- find checklists for fish, birds, mammals to compare with -->

>>>>

## Results and Discussion

@@

We retrieved `r format(n_distinct(combined_species$worms_valid_name), big.mark = ",")` species (`r n_distinct(na.omit(combined_species$worms_phylum))` phyla) of marine invertebrates in the continental US EEZ by combining  `r format(nrow(idigbio_us_inverts_records_species), big.mark=",")` museum records from iDigBio (`r format(n_distinct(idigbio_us_inverts_records_species$worms_valid_name), big.mark = ",")` species, `r n_distinct(na.omit(idigbio_us_inverts_records_species$worms_phylum))` phyla), and `r format(nrow(obis_us_inverts_records_species), big.mark=",")` records from OBIS (`r format(n_distinct(obis_us_inverts_records_species$worms_valid_name), big.mark=",")` species,
`r n_distinct(na.omit(obis_us_inverts_records_species$worms_phylum))` phyla).

@@

Marine invertebrates are underrepresented in iDigBio and OBIS. This underrepresentation is manifest in the small fraction of records that marine invertebrates represent. Extant marine invertebrates occurring in the US EEZ represent  `r format_output(nrow(idigbio_us_inverts_records_species)/idigbio_stats$total_no_fossil*100)`% of the total number of non-fossil iDigBio records (`r round(idigbio_stats$total_no_fossil/1e6, 1)` million), and `r format_output(nrow(idigbio_us_inverts_records_species)/idigbio_stats$usa_no_fossil*100)`% of the US records (`r round(idigbio_stats$usa_no_fossil/1e6, 1)` million). Marine invertebrates however represent between 11% and 18% of the described global diversity (number of described marine inverterbrates species: 200,000-210,000 [@Appeltans2012], total number of described species: 1.2-1.9 million [@Mora2011,@Chapman2009]).

We compared the total number of species represented in iDigBio and OBIS for each major division of life to the total described diversity based on the WoRMS estimates[@worms-stats]. While marine vertebrates represent  `r total_vert_prop`% of the global marine diversity, they represent `r idigbio_vert_prop`% and `r obis_vert_prop`% of the species in iDigBio and OBIS respectively (Fig. \@ref(fig:kingdom-diversity-plot)). This underrepresentation is also apparent in the collecting effort: the median number of specimens in iDigBio per species ranges between `r kng_medians %>% filter(sub_kingdom=="animalia") %>% pull(rg)` for marine invertebrate phyla, but is `r kng_medians %>% filter(sub_kingdom=="animalia - vertebrates") %>% pull(rg)` for marine vertebrates.

```{r itis-stats}
prop_molluscs <- itis_molluscs_in_dbs %>%
    dplyr::summarize(
               prop_idigbio_us = format_output(mean(n_idigbio_geo_us > 0,
                                                    na.rm = TRUE)* 100),
               prop_obis_us = format_output(mean(n_obis_geo_us > 0,
                                                 na.rm = TRUE) * 100)
           )

prop_crustaceans <- itis_crustaceans_in_dbs %>%
    dplyr::summarize(
             prop_idigbio_us = format_output(mean(n_idigbio_geo_us > 0,
                                                  na.rm = TRUE) * 100),
             prop_obis_us = format_output(mean(n_obis_geo_us > 0,
                                             na.rm = TRUE) * 100)               
           )
```

To assess the fraction of digitization of the fauna, we compared the species included in iDigBio and OBIS with the comprehensive checklists for molluscs [@turgeon1998] and crustaceans [@mclaughlin2005] that are available for the United States. For molluscs, of the `r format(nrow(itis_molluscs_in_dbs), big.mark=",")` species included in the checklist, `r prop_molluscs$prop_idigbio_us`% are in iDigBio and `r prop_molluscs$prop_obis_us`% are in OBIS. For crustaceans, of the `r format(nrow(itis_crustaceans_in_dbs), big.mark=",")` species included in the checklist, `r prop_crustaceans$prop_idigbio_us`% are in iDigBio and `r prop_crustaceans$prop_obis_us`% are in OBIS.
